---
title: "Illustrative Pollen Diagram"
author: "Andrew Lane"
date: "2 April 2020"
output: html_document
---

## Introduction

In this notebook we plot a pollen diagram using the `rioja` package. The objective is to use a selection of pollen species for a specific study site to illustrate what a pollen diagram looks like in general, and to show what the signal of arrival of human agriculture might look like as a signal. Use the Navarres example which Carrion et al. 2010 show in their paper as inspiration.


## Package to use

See documentation for `rioja` [here][rioja-docs] and a blog post illustrating its use [here][rioja-blog].

[rioja-docs]: https://cran.r-project.org/web/packages/rioja/rioja.pdf
[rioja-blog]: https://www.benjaminbell.co.uk/2018/02/pollen-diagrams-in-r-using-rioja-part-1.html

Install `rioja` if not already installed
```{r install-rioja, results="hide"}
if (!require("rioja", quietly = TRUE)) {
  install.packages("rioja")
  require("rioja")
}
```

## Load data

We use data prepared in the Jupyter notebook `make_pollen_abundance_timeseries.ipynb`. The data files for each study site are found in the file `../outputs/<sitecode>/selected_species_pol_pct.csv`.

```{r}
setwd('/home/andrew/AgroSuccess/agrosuccess-data/pollen-abundance')
nav_data <- read.csv('../outputs/navarres/selected_species_pol_pct.csv')
```

Separate data from age before present tick labels
```{r}
earliest_date <- 12000
nav_data <- nav_data[which(nav_data[, 1] < earliest_date), ]

agebp <- nav_data$agebp
nav_data$agebp <- NULL
nav_data$Plantago <- NULL # Remove Plantago, all 0
```

```{r}
max(nav_data$Pinus)
```


```{r}
# rioja::strat.plot(ma.pollen1, yvar=y.scale, y.tks=y.scale, y.rev=TRUE, plot.line=FALSE, plot.poly=FALSE, plot.bar=TRUE, col.bar=p.col, lwd.bar=10, scale.percent=TRUE, xSpace=0.01, x.pc.lab=TRUE, x.pc.omit0=TRUE, las=2)

rioja::strat.plot(
  nav_data,
  yvar = agebp,
  # y.tks = agebp,
  y.rev = TRUE,
  scale.minmax = TRUE,
  plot.line=TRUE
  ,
  plot.poly=TRUE,
  plot.bar=FALSE,
  # col.bar=p.col,
  scale.percent=FALSE,
  xSpace=0.02,
  x.pc.lab=TRUE,
  x.pc.omit0=TRUE,
  las=2
)
```

## Work still to do

1. Lower bound for pinus is wrong
2. Vertical axis labels are cropped
3. Pinus pinaster's name has been changed and is cropped at top of plot
4. Vertical axis values don't increase linearly. To solve this would involve reindexing the `agebp` values so they increase linearly and interpolating missing values. For the interpolation we can use [`imputeTS::na_interpolation`](https://www.rdocumentation.org/packages/imputeTS/versions/3.0/topics/na_interpolation), but I can't see a straight-forward way to do the reindexing.
5. Plot a horizontal line at 6000 BP to indicate the arrival of local Neolithic settlement and the beginning of burning.
6. Add colours to plots
